---
title: "P8105 Homework 2"
author: "Junxian Chen ( jc5314 )"
date: "9/27/2019"
output: github_document
---

# Problem 1

### Part 1. Read and clean the Mr. Trash Wheel sheet

Firstly, load nessary libraries:

```{r, message = FALSE}
library(tidyverse)
library(readxl)
```

Next, read in the data as `trash_wheel_data` and do the clean:

```{r}
file_path = "./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx"

trash_wheel_data = 
  read_excel(file_path, sheet = 'Mr. Trash Wheel', range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = as.integer(round(sports_balls))
  )
```

### Part 2. Read and clean precipitation data for 2017 and 2018

```{r}
# Read and clean precipitation data for 2017

pre_2017_data = 
  read_excel(file_path, sheet = '2017 Precipitation', skip = 1) %>%
  janitor::clean_names() %>% 
  drop_na(total)  %>% 
  mutate(year = 2017)
```

```{r}
# Read and clean precipitation data for 2018

pre_2018_data = 
  read_excel(file_path, sheet = '2018 Precipitation', skip = 1) %>%
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = 2018)
```

```{r, message = FALSE}
# Combine precipitation datasets and convert month to a character variable

pre_2017_18 = 
  full_join(pre_2017_data, pre_2018_data) %>%
  mutate(month = month.name[month]) %>%
  replace_na(list(month = 'Sum'))
```

### Part 3. A brief summary about these data


# Problem 2

```{r, message = FALSE}
# read in and clean the data from pols-month.csv

pols_data = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>%
  separate(mon, c('year', 'month', 'day'), convert = TRUE) %>%
  mutate(
    month = month.name[month],
    president = ifelse(prez_gop == 0, 'dem', 'gop'),
  ) %>%
  select(-c(day, prez_dem, prez_gop))
```

(Note: Because the `prez_gop` data from '1974/08/15- 1974/12/15' is recorded as `2` instead of `0` or `1`, I used `ifelse(prez_gop == 0, 'dem', 'gop')` in the code above to avoid dealing with `2` in the classification.)

```{r, message = FALSE}
# read in and clean the data from snp.csv

snp_data = 
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>%
  separate(date, c('month', 'day', 'year'), convert = TRUE) %>%
  mutate(month = month.name[month]) %>%
  select(year, month, close)
```

```{r, message = FALSE}
# read in and clean the data from unemployment.csv

unemployment_data = 
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>%
  pivot_longer(
    Jan:Dec,
    names_to = 'month',
    values_to = 'unemployment rate'
  ) %>%
  mutate(
    month = month.name[match(month, month.abb)]
  ) %>%
  janitor::clean_names()
```

Merge dataset`snp` into `pols`:

```{r}
pols_snp_merge = full_join(pols_data, snp_data, by = c('year', 'month'))
```

Then merge `unemployment` into the results:

```{r}
data_merge = full_join(pols_snp_merge, unemployment_data, by = c('year', 'month'))
```

### A brief summary about these data


# Problem 3

### Part 1. Load and tidy the data

```{r, message = FALSE}
baby_name_data = 
  read_csv("./data/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  distinct() %>%
  mutate(
    childs_first_name = str_to_title(childs_first_name),
    gender = str_to_title(gender),
    ethnicity = 
      str_to_title(
        ifelse(ethnicity ==  'BLACK NON HISP', 'BLACK NON HISPANIC',
          ifelse(ethnicity ==  'WHITE NON HISP', 'WHITE NON HISPANIC',
            ifelse(ethnicity == 'ASIAN AND PACI', 'ASIAN AND PACIFIC ISLANDER', ethnicity
            )
          )
        )
      )
  )
```

### Part 2. A table showing the rank in popularity of the name “Olivia” over time

```{r}
olivia_popularity = 
  baby_name_data %>%
  arrange(year_of_birth) %>%
  filter(childs_first_name == 'Olivia' & gender == 'Female') %>%
  select('Baby`s Name' = childs_first_name , year_of_birth, 'Sex of the Baby' = gender, 
         'Mother`s Ethnicity' = ethnicity, rank) %>%
  pivot_wider(
    names_from = 'year_of_birth',
    values_from = 'rank'
  ) %>%
  rename('Rank in 2016' = '2016', 'Rank in 2015' = '2015', 'Rank in 2014' = '2014', 
         'Rank in 2013' = '2013', 'Rank in 2012' = '2012', 'Rank in 2011' = '2011')

# show the table

print.data.frame(olivia_popularity)
```

### Part 3. A table showing the most popular name among male children over time
```{r}
popular_name_male = 
  baby_name_data %>%
  arrange(year_of_birth) %>%
  filter(gender == 'Male' & rank == '1') %>%
  select(childs_first_name, year_of_birth, 'Mother`s Ethnicity' = ethnicity) %>%
  pivot_wider(
    names_from = 'year_of_birth',
    values_from = 'childs_first_name'
  ) %>%
  rename('Most pop name in 2016' = '2016', 'Most pop name in 2015' = '2015', 'Most pop name in 2014' = '2014', 
         'Most pop name in 2013' = '2013', 'Most pop name in 2012' = '2012', 'Most pop name in 2011' = '2011')

# show the table

print.data.frame(popular_name_male)
```

### Part 4. A scatter plot showing the number of children with a name against the rank in popularity

```{r, fig.width = 10}
names_poplarity_plot =
  baby_name_data %>%
  filter(gender == 'Male', ethnicity == 'White Non Hispanic', year_of_birth == '2016') %>%
  ggplot(aes(x = rank, y = count)) +
  geom_point() +
  labs(x = 'Rank in popularity of the name', 
       y = 'Number of children with the name',
       title = 'Name poplularity and counts for male, white non-hispanic children born in 2016') 

names_poplarity_plot
```

